---
title: "412 Project - Seurat Processing"
author: "Kristi Lin-Rahardja"
date: "5/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup .{tabset}

## Libraries

```{r setup, include=FALSE}
library(useful); library(devtools)
library(Seurat); library(Matrix); library(clustree)
library(tidyverse); library(dplyr)
library(ggplot2); library(ggridges); library(gridExtra)
```

## Load data

Set the full path to the main directory 
```{r set wd path}
rm(list=ls())
dir <- "/Users/link7/Documents/412/412_proj/"
```

Load data
```{r load data}
raw.data <- read.csv(paste(dir,"csv_files/S01_datafinal.csv", sep=""), header=T, row.names = 1)
dim(raw.data)
head(colnames(raw.data))
```

Load metadata
```{r load metadata}
metadata <- read.csv(paste(dir,"csv_files/S01_metacells.csv", sep=""), row.names=1, header=T)
head(metadata)
```

Save both objects as RData object for quicker loading 
```{r save object}
save(list=c("metadata","raw.data","dir"), file=paste(dir,"objects/S01_Data_and_metadata.RData", sep=""))
```


# Data Preparation .{tabset}

## Create Seurat Object

load and clean raw data
```{r}
osi.raw.data <- read.csv(paste(dir,"csv_files/neo-osi_rawdata.csv", sep=""), row.names = 1)
colnames(osi.raw.data) <- gsub("_S.*.homo", "", colnames(osi.raw.data))
head(osi.raw.data)
tail(osi.raw.data)
# drop sequencing details from gene count table
row.names(osi.raw.data)[grep("__", row.names(osi.raw.data))]
osi.raw.data <- osi.raw.data[-grep("__", row.names(osi.raw.data)),]
```

Make osi.metadata by cell from osi.metadata by plate
```{r}
osi.metadata <- read.csv(paste(dir, "csv_files/neo-osi_metadata.csv", sep = ""))
osi.meta.cell <- as.data.frame(colnames(osi.raw.data))
osi.meta.cell <- data.frame(do.call('rbind', strsplit(as.character(osi.meta.cell$`colnames(osi.raw.data)`),'_',fixed=TRUE)))
rownames(osi.meta.cell) <- paste(osi.meta.cell$X1, osi.meta.cell$X2, sep = "_")
colnames(osi.meta.cell) <- c("well", "plate")
osi.meta.cell$cell_id <- rownames(osi.meta.cell)
osi.metadata <- left_join(osi.meta.cell, osi.metadata, by = "plate")
rownames(osi.metadata) <- osi.metadata$cell_id
head(osi.metadata)
unique(osi.metadata$plate)
```

Find ERCC's, compute the percent ERCC, and drop them from the raw data.
```{r}
erccs <- grep(pattern = "^ERCC-", x = rownames(x = osi.raw.data), value = TRUE)
percent.ercc <- Matrix::colSums(osi.raw.data[erccs, ])/Matrix::colSums(osi.raw.data)
ercc.index <- grep(pattern = "^ERCC-", x = rownames(x = osi.raw.data), value = FALSE)
osi.raw.data <- osi.raw.data[-ercc.index,]
dim(osi.raw.data)
```

Create the Seurat object with all the data (unfiltered)
```{r}
osi_object <- CreateSeuratObject(counts = osi.raw.data)
osi_object <- AddMetaData(object = osi_object, metadata = osi.metadata)
osi_object <- AddMetaData(object = osi_object, percent.ercc, col.name = "percent.ercc")
# Changing nUMI column name to nReads
colnames(osi_object@meta.data)[colnames(osi_object@meta.data) == 'nUMI'] <- 'nReads'
head(osi_object@meta.data)
```


Filter cells so that remaining cells have nGenes >= 500 and nReads >= 50000
```{r}
osi_object_filtered <- subset(x=osi_object, subset = nCount_RNA > 50000 & nFeature_RNA > 500)
osi_object_filtered
osi_object_filtered
```

```{r}
VlnPlot(osi_object_filtered, features = "nFeature_RNA")
VlnPlot(osi_object_filtered, features = "nCount_RNA", log = TRUE)
```



# Subset .{tabset}

## Clean

```{r}
load(file=paste(dir,"objects/S03_Merged_main_filtered_with_neo_osi.RData", sep=""))
```

Drop any samples with 10 or less cells
```{r}
main_tiss_filtered1@meta.data$sample_name <- as.character(main_tiss_filtered1@meta.data$sample_name)
sample_name <- as.character(main_tiss_filtered1@meta.data$sample_name)
# Make table 
tab.1 <- table(main_tiss_filtered1@meta.data$sample_name) 
# Which samples have less than 10 cells 
samples.keep <- names(which(tab.1 > 10))
metadata_keep <- filter(main_tiss_filtered1@meta.data, sample_name %in% samples.keep)
# Subset Seurat object 
tiss_subset <- subset(main_tiss_filtered1, cells=as.character(metadata_keep$cell_id))
tiss_subset
```

Check sample and patient counts
```{r}
table(tiss_subset@meta.data$sample_name)
table(tiss_subset@meta.data$patient_id)
```

Normalize and scale the data
```{r}
tiss_subset <- NormalizeData(object = tiss_subset, scale.factor = 1e6)
tiss_subset <- ScaleData(object = tiss_subset)
```

Find variable genes
```{r}
tiss_subset <- FindVariableFeatures(object = tiss_subset)
```

# Dimensionality Reduction .{tabset}

## PCA

Perform PCA
```{r}
tiss_subset <- RunPCA(object = tiss_subset, do.print = FALSE)
```

Visualize variance along each component
```{r}
ElbowPlot(tiss_subset)
```

Visualize first two PCs
```{r}
DimPlot(tiss_subset, reduction = 'pca')
```

Visualize top genes and their loadings 
```{r}
VizDimLoadings(tiss_subset, dims = 1:4, reduction = "pca", nfeatures = 10)
```

Visualize top genes in principal components
```{r}
DimHeatmap(tiss_subset, dims = 1:7, cells = 100, balanced = TRUE)
```

Set nPCs from above visualizations
```{r}
n.pcs <- 20
```

Construct Neighbor graph
```{r}
tiss_subset <- FindNeighbors(object = tiss_subset, dims = 1:20, verbose = T)
```

Find Clusters 
```{r}
tiss_subset <- FindClusters(object = tiss_subset, verbose = T, resolution = 0.5)
```

## tSNE

Run and project TSNEs
```{r}
tiss_subset <- RunTSNE(tiss_subset, dims = 1:15)
DimPlot(tiss_subset, reduction = "tsne", label = TRUE)
```

Construct Neighbor graph 
```{r}
tiss_subset <- FindNeighbors(object = tiss_subset, dims = 1:20, verbose = T)
```

Find Clusters 
```{r}
tiss_subset <- FindClusters(object = tiss_subset, verbose = T, resolution = 0.5)
```




# Marker Expression Analysis .{tabset}

## Visualize

```{r}
#tsne + pca showing how tx history spreads across clusters
p = TSNEPlot(object = tiss_subset, group.by = 'treatment_history')
ggsave(paste(dir,"plot_out/mine/tsne_tx.png", sep=""), p, width = 10, height = 10)

p = DimPlot(tiss_subset, reduction = 'pca', group.by = "treatment_history")
ggsave(paste(dir,"plot_out/mine/pca_tx.png", sep=""), p, width = 10, height = 10)


#define what markers we want to observe
features <- c(#"MDM2", "CCND1", "MTOR")  #survival/proliferation pathways
              #"TLR2", "TLR4", "IRF3") #TLR pathway; proinflammation
              #"BAD", "BID", "BAX", "FADD", "CRADD", "APAF1") #apoptosis markers https://www.rndsystems.com/pathways/apoptosis-signaling-pathway
              #"CD27", "CD40", "CD28", "ICOS") #immune stimulatory checkpoint molecules
              "CTLA4", "BTLA", "LAG3", "SIGLEC7") #inhibitory checkpoint molecules
for (i in 1:length(features)){ #loop through the features to make the next plots individually per gene
  #tsne highlighting expression of certain markers
  p=FeaturePlot(tiss_subset, features = features[i], split.by = "treatment_history")
  ggsave(paste(dir,"plot_out/mine/", features[i], "_featplot.png", sep=""), p, width = 15, height = 7)

  
  #expression in clusters, comparing naive vs prior tx
  p=RidgePlot(tiss_subset, features = features[i], group.by= "treatment_history")
  ggsave(paste(dir,"plot_out/mine/", features[i], "_tx_ridge.png", sep=""), p, width = 15, height = 7)
  
  VlnPlot(tiss_subset, features = features[i], group.by= "treatment_history")#,split.by = "treatment_history")
  ggsave(paste(dir,"plot_out/mine/", features[i], "_tx_vln.png", sep=""), p, width = 15, height = 7)
}
```
